---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# plotPlotR

<!-- badges: start -->

<!-- badges: end -->

The Agricultural research and management operations in farms are increasingly becoming digital. Depending on the applications, the equipment used in farm operations and data collection are now
increasingly connected to the internet, satellites and generates rich meta-data.

**plotPlotR** referring to plot plotter is a package containing set of functions to create plots on a field trial as segments or boundaries in a field experiment.

The functions are based on the implementation of "**sf**" package in R and a takes syntax of dplyr and magrittr packages. The package provides set of functions that are simple for R users to create or manipulate plot boundaries. Users who may not be familiar with spatial data or want functions that can be used easily, repeatedly can use this package.

## Installation

You can install the development version of plotPlotR from [GitHub](https://github.com/sin17m/plotplotr) with:

``` r
# install.packages("devtools")
devtools::install_github("sin17m/plotplotr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(plotPlotR)
## basic example code

cornersData # a data frame with 4 XY points referring to the four corners of an experiment

dat_unit_m <- wgs84_to_unitM(cornersData, 28355) # converting to unit M CRS
#Above step is optional but recommended

dat_plots <- corners_to_plots(dat_unit_m, 80, 24.2, 11, 20)

dat_RR <- addRunRange(dat_plots, "BL")

plot(dat_RR)



```
