---
title: "plotPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plotPlotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

`# install.packages("devtools")`

`devtools::install_github("AgDataNomad/plotPlotR")`

## Library

```{r setup}
library(plotPlotR)
```

## Introduction

**plotPlotR** package is developed to provide R users involved in field experiments to create or manipulate geo-spatial plot layouts for their experiments. This package is small and has a handful of functions to perform QGIS like field plot creation and manipulation which then can be used to extract values from different types of raster or adding values to plots and create shapefiles etc.

This package is particularly useful for users when working on raster data generated from Unmanned Aerial Vehicles (drones, UAVs) or if users want to associate non-spatial data to the plot data.

The core of plotPlotR is the creation of plot segments. If you have obtained corner points from [Google Maps](https://www.google.com/maps), [Google Earth](https://earth.google.com/web/), GPS sensors on farm machine or from a GPS survey tool, then you will be able to create plots easily using the functions in this package. You can also assign Run and Range values to your plots which then can then be moved both vertically and horizontally to account for any discrepancies.

This package provide user friendly functions to deal with field plot creation and manipulation.

## Good to know

It is strongly recommended that you use a CRS that are in meters. Function `read_input_dat()` is a way users read data for futher processing using this package. This function will help (enforce) users to assign and change CRS of the input data. You may choose to use [epsg.io](https://epsg.io/) to find the appropriate CRS for your region. The examples in this package uses CRS [28355](https://epsg.io/28355), as we are based in Australia.

-   Be familiar with the terminology used in this package. People who are not familiar with field operations may find the terminology used for arguments in functions new. Please refer to functions help menu or [reference](https://agdatanomad.github.io/plotPlotR/reference/index.html) for more information. Below examples contains some terminology you may not be familiar with but are used extensively in this package.

    -   `Run` is plots along tractor runs
    -   `Range` is the other axis
    -   `exp_length` refers to length of your experiment (length of Runs)
    -   `exp_width` refers to width of your experiment

-   The package doesn't call user to provide precise direction in North, South, East, West in most case. This is because, field experiments are not in aligned to any one direction and some users may not be able to precisely capture and record that information. Lot of functions in this package uses relative terms like, up, down, left, right, top and bottom. We believe this is easier for users to understand and use. Please refer to the documentation of functions where those relative terms are used.

## Experiment layout

An experiment layout for the purposes of this package is a rectangular grid layout, where each smaller rectangles are plots. The plots can be represented by its position on two axis, `Run` and `Range`. Plots along the direction of the tractor travel are called `Run` and plots on the other axis are called `Range`.

The experiments themselves has experiment length `exp_length` which is akin to the length of a `Run` and experiment width `exp_width` referring to the length of a `Range`. `exp_length` is measure plot length with gap multiplied by the number of ranges and `exp_width` is tractor wheel width multiplied by the number of runs. Each run has the maximum number of ranges and vice versa.

For several practical reasons, in most cases, experiment length are typically longer than experiment width but good not to make assumptions. There may be some cases when your experiment is part of a larger experiment, you many fewer ranges than runs, in that case your experiment length might be smaller that experiment width.

Run and Range may not the terminology you use. Rows, Runs, Row, Run, rowNo, runNo, row_Num, run_NUm are aliases for Runs that we have encountered. Similarly for Range, column, Column, col etc are used as alias. It is hard to deal with so many aliases, so try to be consistent and where needed provide what alias is being used in your experiment.

```{r Experiment layout example}

dat <- read_input_dat(cornersData, 4326, 28355)

corners_to_plots(sf_object = dat, #package includes cornersData 
                 exp_length = 80, #number of ranges multiplied by plot length including gaps
                 exp_width = 24.2, #number of runs multiplied by tractor wheel width
                 n_runs = 11, #total number of runs
                 n_ranges = 20) #total number of ranges

```

## Getting started

### Reading data

The function `read_input_dat()` is used to bring your coordinates data into the plotPlotR environment. `read_input_dat` allows users to input a csv file path or a dataframe loaded in the `R` environment or any common shape file formats.

*plotPlotR* requires all shapefiles to have a Coordinate Reference System (CRS) and more importantly it requires coordinates in unit meters to compute different functions. Coordinate Reference System is a coordinates based system to locate geographical entities and is the core principles of GIS workflows. Most common CRS is WGS84 or EPSG 4326, which uses longitudes and latitudes to locate geographical entities. This CRS is what gets used in applications such as navigation systems, google maps, drone navigation etc. Orthomosaics generated using a stitching software often has CRS 4326.

When reading data from a CSV file or from a dataframe, users need to supply the X and Y values using `XY` argument and the CRS using the `input_CRS` arg. `input_CRS` argument takes any default CRS formats, just entering the EPSG code as a number works, for example `input_CRS = 4326`. `XY` argument is not required if and when the csv file or the dataframe has XY as the first two columns.

Here is an example of reading data using the **cornersData** dataset included in the package. `read_input_dat` function is used without declaring `to_CRS_M` argument. NOTE: When `to_CRS_M` is not supplied, plotPlotR automatically assigns a **WGS 84 / Pseudo-Mercator** CRS.

```{r reading_dat}

dat <- read_input_dat(input_dat = cornersData,
                      input_CRS = 4326)

dat

```

Here is an example of reading data with a `to_CRS_M` agrument.

```{r}

dat <- read_input_dat(input_dat = cornersData,
                      input_CRS = 4326,
                      to_CRS_M = 28355)

dat

```

An example of reading data from a dataframe

```{r}

dat <- cornersData

class(dat) ##dataframe

df_dat <- read_input_dat(input_dat = dat, input_CRS = 4326)

```

Here is an example of reading input data when the dataframe or the CSV file doesn't contain X and Y as the first two columns. `read_input_dat` function will convert the XY columns to *geometry* and preserves all other columns.

```{r}

#rearranging and renaming columns and adding a new column to demonstrate read_input_dat function

dat <- dat[,c(3,1,2)]

colnames(dat) <- c("A","B","C")

dat$comment <- "some_gps_data"

df_dat <- read_input_dat(input_dat = dat,
                         input_CRS = 4326,
                         XY = c("B", "C"), # columns names of X and Y coordinates
                         to_CRS_M = 28355)

df_dat


```

### Writing data

Writing outputs of **plotPlotR** is done through calling `write_output_obj()` function. This function takes two arguments; `output_dat` and `file_name`. File name is the file path with the desired extention type. For example *geoJSON*, *kml* etc. 

```{r}

dat <- read_input_dat(input_dat = cornersData,
                      input_CRS = 4326)

#write_output_obj(dat, "path_to_folder/file_name.geoJSON")

```


## Creating plot layout

Creating plot layout is the core function of the **plotPlotR** package and it can be done in one of two ways depending on what information is available.

1. From 4 coordinates or sets of 4 coordinates representing the corners of the experiment with center of corners being the center of the experiment. Function`corners_to_plots` is used to generate plot layouts from the corner points. The corner points, themselves can be generated using different ways, if you have a raster you can use the function `collect_corners` to get corner points. Users can also use services such as google earth, QGIS, Ground Control Points data to get 4 corner points. 
2. From trip points generated from tractors/seeders etc using the `make_plots` function. Modern tractors uses GPS or similar systems to drive on a paddock to sow seeds etc. These trip points, one for each plot can be used to create a field layout. Variation of make plots, is splits plots, where two or three varieties are sown in each plot. Users can use `make_split_plots` function to generate split plots from one trip point or from the plot layout generated using `corners_to_plots`

### Corners to plots

`corners_to_plots` uses corners data provided to calculate the center and the orientation of the experiment. From the exp_length, exp_width, n_runs, n_ranges arguments a rectangular grid with the specified number of Runs and Ranges are generated. 

Optional `exp_orientation` sets the orientation of the experiment, Runs can be aligned North-South (vertically) or East-West (horizontally) in the rectangular grid. Sizes of plots generated can be customised using the optional `plot_length` and `plot_width` argument.

NOTE: Orientation of the exp is correlated to experiment length `exp_length`, the orientation of the `Run`, length of the plots. If the orientation of the experiment is in **NS**, then the `exp_length`, `plot_length` and the `Run` are in the same direction.

```{r}

plots_data <- corners_to_plots(sf_object = dat,
                               exp_length = 80,
                               exp_width = 40,
                               n_runs = 2,
                               n_ranges = 4)

plot(plots_data)

```


Changing the orientation of the experiment to **EW** East-West using the optional `exp_orientation` argument

```{r}

plots_data <- corners_to_plots(sf_object = dat,
                               exp_length = 80,
                               exp_width = 20,
                               n_runs = 2,
                               n_ranges = 4,
                               exp_orientation = "EW")

plot(plots_data)

```

Specifying plot length and plot width using the OPTIONAL `plot_length` and `plot_width` argument

```{r}

plots_data <- corners_to_plots(sf_object = dat,
                               exp_length = 80,
                               exp_width = 40,
                               n_runs = 2,
                               n_ranges = 4,
                               plot_length = 15,
                               plot_width = 8.6)

plot(plots_data)

```

### Points to plots

If you have coordinates (points) for each plot in the experiment, say from a seeder or other device, the function `make_plots()` can be used to create plot layouts of the experiment. `make_plots` function takes coordinates as points (a shape file or an sf object with POINT geometry) as input and using the `plot_length()` and `plot_width()` arguments generates the plot layout.

In the below example, I am using `cornersData` as coordinates of 4 plots. Then using `make_plots()` function, I am generating a plot layout.   

```{r}

plot_layout <- make_plots(points_data = dat, 
                          plot_length = 75, 
                          plot_width = 20)

plot(plot_layout)

```

### Split plots

Split plots are type of field experiments where more than one genotype/variety is sown at each plot. `make_split_plots()` function allows users to create a split plot layout from plot center points or from plot layouts generated using `corners_to_plots`. 

```{r}

oneplot <- corners_to_plots(dat, 
                            exp_length = 80, 
                            exp_width = 30, 
                            n_runs = 5,
                            n_ranges = 8)

split_plot_layout <- make_split_plots(oneplot,
                                      split_dist = 2,
                                      split_into = 2,
                                      plot_length = 7.5,
                                      plot_width = 2)

plot(split_plot_layout)

```

## Plot manipulation

**plotPlotR** provides functions to transform experimental layouts in a few different ways. For easier transformation, plots needs to be identified by its coordinates, i.e. each plot get assigned values of Run and Range. Runs and Ranges are a common local coordinate reference system used in field experiment to identify a plot. Runs and Ranges are unique to each experiment and has no absolute reference to the experiment spatially. This simple system however is quite useful and adequate for people working on field experiments.  

### Adding Run and Range values

`addRunRange()` function adds Run and Range values to the plots in the experiment. For assigning Run and Range values, users need to provide three pieces of information. The number of Runs using the `n_runs` argument, number of Ranges using the `n_ranges` argument and the position of the Row 1 and Range 1 using the `run1_range1` argument. The argument `run1_range1` takes one of 4 values from "TR", "TL", "BR" and "BL" which represents top right, top left, bottom right and bottom left respectively. 

*NOTE: The southern end is the bottom of an experiment in North-South direction and similarly the eastern end is the bottom of the experiment in the East-West direction. Assuming someone standing at the bottom of their experiment looking at the top end, their left side is the left side of the experiment and similarly their right side is the right side of the experiment. *

```{r}

plots_data <- addRunRange(plots_data, 
                          run1_range1 = "BL",
                          n_runs = 2,
                          n_ranges = 4)

plot(plots_data)

```


Lets explore few transformations

### Rotating the whole experiment layout

`rot_plots()` functions takes a sf-object and rotate on its center by the value provided in the `rotation_angle` argument. Users can rotate the whole experiment or just the plots using the `selection` option. Rotation angle is declared in degrees and can be positive or negative value.  

```{r}

plot(plots_data)

rotated_exp <- rot_plots(sf_object = plots_data, rotation_angle = 90, selection = "experiment")

plot(rotated_exp)

```

### Rotating individual plots

Rotating individual plots is similar to rotating the entire experiment, only difference being the orientation of the experiment will be the same and individuals plots were rotated on its axis. 

```{r}

plot(plots_data)

rotated_plots <- rot_plots(sf_object = plots_data, rotation_angle = -10, selection = "plots")

plot(rotated_plots)

```

### Plot displacement

Plots (or experiments as a whole) can be moved horizontally, vertically, along its run or range or along the compass direction in a number of ways using functions in **plotPlotR**. 

Moving runs can be done using the `movePlotRuns()` function. Movement of plots are done through two arguments, `moveDirection` and `moveDistane`. `moveDirection` takes "left", "right", "up" or "down" and `moveDistance` takes a distance value in meters. There is a optional argument `runSelection`, which when not supplied applies the movement to all plots.

```{r}

moved_run <- movePlotRuns(plots_data, moveDirection = "up", moveDistance = 5, runSelection = 2)

plot(moved_run)

```

Similarly plots can be moved range wise. By default when `rangeSelection` is not supplied, all Ranges are moved. 

```{r}

moved_range <- movePlotRanges(plots_data, moveDirection = "right", moveDistance = 5, rangeSelection = 2)

plot(moved_range)

```

Entire experimental layout can be moved along the north, south, west or east direction using the `moveAllPlots()` function. This function takes distance to move the experimental layout horizontally and vertically. `xy_move` is a vector of length two, first element is the distance to move in X direction and the second element is the distance to move in the Y direction. Distances can be positive or negative values. 

```{r}

moveAllPlots(plots_data, xy_move = c(2,4))

```

### Plot resizing

Sizes of plots in an experiment can be manipulated using the function `make_plots()`. This function usually takes points data and produce an experimental layout but can also be used to modify plot dimensions. Users can supply `plot_length` and `plot_width` arguments.

```{r}

modified_plot_dimensions <- make_plots(plots_data, 
                                       plot_length = 15, 
                                       plot_width = 15)

#generating square plots with dimension 15 meters by 15 meters

plot(modified_plot_dimensions)

```

